---
title: "Linear & Logistic Regression - Answers"
author: "Samantha-Jo Caetano"
date: "October 4, 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Let's work through the following R code:

```{r}
library(tidyverse)
```


# Loading in Data


The `dpylr` package has some data about Star Wars characters. Let's assume it is a representative sample of all characters seen in Episodes 1 to 9.


```{r, message=FALSE}

starwars<-starwars

glimpse(starwars)

head(starwars)

```

Let's start by playing around with some SRS ratio estimation. Use the `svyglm()` function in the `survey` library to run a ratio estimation of mass given height. Assume it was a Simple Random Sample and use the finite population correction with $N=200$.

```{r}
#install.packages("survey")
library(survey)
## Using the Survey Library
N=200
n=nrow(starwars)

my.design <- svydesign(id=~1,
                       data=starwars,
                       fpc=rep(N, n))
  
svyratio(~mass, ~height, design=my.design)

svyratio(~mass, ~height, design=my.design, na.rm=TRUE)

```


Let's try taking this a step further playing around with some Stratified random sampling ratio estimation. Use the `svyglm()` function in the `survey` library to run a ratio estimation of mass given height. Assume the study wished to stratify by gender, use the finite population correction with $N=200$ and assuming 20% of the population are "feminine" and 80% are masculine.

```{r}
#install.packages("survey")
#library(survey)
## Using the Survey Library
N=200
n_masc=nrow(starwars %>% filter(gender=="masculine"))
n_fem=nrow(starwars %>% filter(gender=="feminine"))

# In the sample there are 17 feminine characters and 66 masculine characters.

## Need to create a weight variable for 25% fem and 75% masculine and then a second variable that spreads the weight amongst the sample.

starwars <- starwars %>% 
  mutate(pop_strata_weight = case_when(gender=="feminine" ~ 40,
                                   gender=="masculine" ~ 160))

starwars <- starwars %>% 
  mutate(samp_strata_weight = case_when(gender=="feminine" ~ 40/17,
                                   gender=="masculine" ~ 160/66))

starwars_noNA <- starwars %>% filter(!is.na(gender))

strata.design <- svydesign(id=~1,
                       strata=~gender,
                       weights = ~samp_strata_weight, 
                       data=starwars_noNA,
                       fpc=~pop_strata_weight)
  
svyratio(~mass, ~height, design=strata.design)

svyratio(~mass, ~height, design=strata.design, na.rm=TRUE)


### If you wish to compare with SRS example
# svyratio(~mass, ~height, design=my.design, na.rm=TRUE)

```



### Task 1: Create a Simple Linear Model to predict mass

Use the `svyglm()` function in the `survey` library to run a linear regression estimation of mass given height. Assume it was a Simple Random Sample and use the finite population correction with $N=200$.

```{r}
#install.packages("survey")
#library(survey)

## Using the Survey Library
n=87
N=200
fpc.srs = rep(N, n)

starwars.design <- svydesign(id=~1, data=starwars, fpc=fpc.srs)

mysvylm <- svyglm(mass ~ height, starwars.design)
summary(mysvylm)
```

What is the standard error of the intercept?

a. -13.81
**b. 17.36**
c. 0.64
d. 0.06


### Task 2: Create a Linear Model to predict mass

Use the `lm()` function in the `survey` library to run a linear regression estimation of mass given height.

```{r}
#install.packages("survey")
library(survey)

## Using the Survey Library
summary(lm(mass ~ height, data=starwars))

```

What is the different between the output here and the outcome in Task 1? What is similar?

**Standard errors are different, but estimates are the same.**

What happens if you change the $N$ in Task 1? Try setting $N=87, 200, 1000, 10000$


### Task 3: Create a Linear Model to predict mass

Run a linear regression estimation of mass given height and sex.

```{r}
## Using the Survey Library
n=87
N=200
fpc.srs = rep(N, n)

starwars.design <- svydesign(id=~1, data=starwars, fpc=fpc.srs)

mysvylm <- svyglm(mass ~ height + sex, starwars.design)
summary(mysvylm)

## Using lm
mymodel<-lm(mass ~ height + sex, data=starwars)
summary(mymodel)

```

Based on the output, what is the estimated mass of a male character who is 175 units tall?

```{r}
-52.85109 + 0.63508*175+20.84051
predict(mysvylm, tibble(height=175, sex="male"))
predict(mymodel, tibble(height=175, sex="male"))
```


### Task 4: Create a Logistic Regression Model to predict if mass exceeds 100lb.

Run a logistic regression model mass being over 100lbs given height and sex. 

```{r}
## Using the Survey Library

## Remove missing mass

starwars <- starwars %>% filter(!is.na(mass))

## Create a new variable called `mass_over100`. 

starwars <- starwars %>%  mutate(
  mass_over100 = case_when(
    mass > 100 ~ 1, ## 1 = Yes, over 100lbs
    mass <= 100 ~ 0)
)


## Using lm
my_glm <- glm(mass_over100 ~ height + sex, data=starwars, family = "binomial")
summary(my_glm)

```


Based off the above output, what is the predicted probability of a male who is 170cm tall being over 100lbs?

**Answer: 0.05729392**

```{r}
pred_odds <- -28.90236+0.0604*170+15.8338*1
exp(pred_odds)/(1+exp(pred_odds))

pred_odds2 <- predict(my_glm, tibble(sex="male", height=170))
exp(pred_odds2)/(1+exp(pred_odds2))

pred_prob <- predict(my_glm, tibble(sex="male", height=170), type = "response")
pred_prob
```


\newpage 

# Logistic Regression UCLA analysis of SRS with population of N=6000 code

```{r}
mydata <- read.csv("https://stats.idre.ucla.edu/stat/data/binary.csv")

## Standard Logistic Regression
mylogit<-glm(admit ~ gre + gpa +
               as.factor(rank), data=mydata, family="binomial")
summary(mylogit)

## log(p/(1-p)) = -3.99 + 0.002*gre + 0.804*gpa-0.675*x3
##                           -1.34*x4 - 1.55*x5


## Survey Estimation for Logistic Regression
n=length(mydata$admit)
N=6000

#install.packages("survey")
library(survey)
## Using the Survey Library
fpc.srs = rep(N, n)

ucla.design <- svydesign(id=~1, data=mydata, fpc=fpc.srs)

mysvyglm <- svyglm(admit ~ gre + gpa + as.factor(rank), 
                   ucla.design, family="binomial")
summary(mysvyglm)

```

